Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-12-06 journal-draft/bare_jrnl_comsoc.tex
modified sec 3

A cluster $C$ is a set of secondary nodes which possess the same set of CCs.
In particular, one cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be composed only by the cluster head.
$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
Cluster size of $C$ is written as $|C|$.
Cluster $C(i)$ means the cluster head of this cluster is $i$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \cap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model and the following problem description are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
$C_i$  & the $i$th potential cluster in a collection of clusters \\
 $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{theorem}
\label{theorem1}
Robust clustering problem in CRN is NP-hard, when the maximum size of clusters is larger than 3, and $\delta_1=1$ and $\delta_2 = N$ .
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.
%This theorem actually indicates that the robust clustering problem in CRN is NP-hard.

Although the robust clustering problem is NP-hard, it can be solved by standard solvers. 
Thus, a central node with all knowledge of all clusters could solve the problem based on your formulation. 
%we need to present this problem in the form of optimization to solve in the centralized manner.

%Binary linear programming problem is NP complement, but with the 
%The result will be used as a comparison with the proposed distributed schemes. %looks for the solution which approaches to the global optimum, and the latter is the solution which is more suitable for the practical implementation.
%
%Exact cover problem can be solved with Knuth's Algorithm X~\cite{dancingLinks_Knuth} as it finds out all the instances of exact cover, then we can choose the one with the biggest sum of weights. 
%To make the formulation possible, we adopt heuristics in both the objective function and the constraints, besides, the objective is not totally aligned with the object in the Definition~\ref{def_centralized_clustering}.
%The problem by the Definition~\ref{def_centralized_clustering} is formulate

@@

\begin{equation}
\begin{aligned}
     &\min\limits_{x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^G (-x_{ij}q_{ij} + (1-w_i)*p) \\
     &\text{subject to}   && \Sigma_{i=1}^G x_{ij} = 1, for \forall j=1, \ldots, N \\
   &&& \Sigma_{j=1}^N x_{ij} = |C_i|*(1-w_i), for \forall i=1, \ldots, G \\
   &&& \text{$x_{ij}$ and $w_i$ are binary variables.}\\
   &&& i\in \{1,2, \cdots G\}, \hspace{0.3cm} j\in \{1,2,\cdots N\}
%\notag
\end{aligned}
\label{centralized_opt}
\end{equation}

The optimization is a binary linear programming problem as shown in~\ref{centralized_opt}.
The objective is to maximize the sum of CCs of all the clusters.
We minimize the opposite of the real objective to make this problem solvable.
%
The variables of this problem are $x_{ij}$ where $i\in \{1,2,\cdots,G-1, G\}$ and $j\in \{1,2,\cdots,N-1, N\}$.
$N$ is the total number of CR users in network $\mathcal{N}$.
$G = |\mathcal{G}|$ and $\mathcal{G}$ is a collective of clusters in a CRN $\mathcal{N}$.
There is $\mathcal{G} \subseteq \mathcal{S}$, where $\mathcal{S}$ is the collection of all the possible clusters as described in Definition~\ref{def_centralized_clustering}.
%In practice, we can solve the problem with $\mathcal{G} = \mathcal{S}$ when $|\mathcal{S}|$ is not large. 
The subscript $i$ indicates the $i$th cluster in $\mathcal{G}$, and $j$ corresponds to the node ID of a CR node. 
$x_{ij}$ indicates whether the CR node $j$ resides in the $i$th potential cluster, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$. 
%
%Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{potential} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
%Potential clusters are the clusters which satisfy the conditions introduced in Section~\ref{sec:model}. 
%Note that the potential clusters include the singleton ones.
%By permitting the existence of the singleton clusters, we can ensure that the $\mathcal{S}'$ in Definition~\ref{def_centralized_clustering} is always be feasible.
Based on the knowledge on $\mathcal{G}$, we construct a $G\times N$ matrix $Q_{G\text{x}N}$ which is shown in Figure~\ref{costant_matrix_Q}. 
Constant $q_{ij}$ is the element of $Q_{G\text{x}N}$, and its subscripts correspond to the $i$th cluster and CR node $j$ respectively.
$q_{ij}= |K(C_i)|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
In other words, each non-zero element $q_{ij}$ means the number of CCs of the cluster $i$ where node $j$ resides.

\begin{figure}[ht!]
\centering
\bordermatrix{~ 		& 1 	& 2 	& 3 	& \cdots & j & \cdots	& N-1 	& N	\cr
                  1 	& |K(C_1)| 	& |K(C_1)| 	& 0 	& \cdots & \cdots &\cdots	& 0 	& 0	\cr
                  2 	& |K(C_2)| 	& 0 	& |K(C_2)| 	& \cdots & \cdots & \cdots 	& 0 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots		& 		& \vdots \cr
				i 	& 0 	& |K(C_i)| 	& 0 	& \cdots  & \cdots & \cdots 	& |K(C_i)| 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots & \cdots & \vdots 		& 		& \vdots \cr
				\vdots 	& 0  	& 0 	& 0 	& \cdots & \cdots & \cdots 	& |K(C_i')| 	& 0	\cr
				G  	& |K(C_G)| & \cdots	 	& 		&  \vdots	& \cdots & \vdots& 		& \vdots \cr}	
\caption{An example of Matrix $Q$, its rows correspond to all possible clusters, and columns correspond to the CR nodes in the CRN. }
\label{costant_matrix_Q}
\end{figure}

Now let us have a look at the second component in the objective function.
%Although the related work SOC~\cite{LIU_TMC11_2} adopts the product of cluster size and number of CCs as its sole metric, SOC still generates a large number of singleton clusters and a few large clusters.
To discourage the optimization algorithm from generating the clusters which don't have the preferred size, the second item compromises the first component in the objective function when those clusters are generated.
$w_i$ is the auxiliary binary variable which denotes whether the $i$th  cluster $C_i$ is chosen by the optimization algorithm or not.

$$
w_i = \left\{ \begin{array}{rl}
0 &\mbox{if $i$th possible cluster $C_i$ is chosen} \\
1 &\mbox{if $i$th possible cluster $C_i$ is not chosen} \\
\end{array} \right.
$$
When $w_i$ is zero, it means $C_i$ is chosen by the optimization algorithm, in this case, the second item in the objective function becomes $p$, which is defined as follows,

$$
p = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$ or $|C_i|=\delta+1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$ or $|C_i|=\delta+2$} \\
\dots
\end{array} \right.
$$
where $\alpha_1, \alpha_2, \cdots$ are positive values, then when $w_i$ is zero ($C_i$ is chosen by the optimization algorithm) and its size is not the desired size $\delta$, the second component will be greater than zero, which will compromise the first component.
In particular, $\alpha$ increases with the divergence between $|C_i|$ and $\delta$ \ie these is $\alpha_2> \alpha_1>0$.
%Because of $w_i$, any cluster which appears in the clustering solution ($w=0$) will compromise the first component.
%Only when that cluster's size is exactly the preferred size $\delta$, the punishment is zero.
%In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
%Then it is easy to notice that when the optimization algorithm adopts singleton clusters, the first xx.
%With such design, the solution is encouraged not to generate singleton clusters.
%doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
%Choice of $\alpha_i$ affects the resultant clusters.

>>>>

We give the definition of cluster in CRN as follows. 
A cluster $C$ is a set of secondary nodes which possess the same set of CCs.
In particular, one cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be composed only by the cluster head.
$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
Cluster size of $C$ is written as $|C|$.
Cluster $C(i)$ means the cluster head of this cluster is $i$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \cap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model and the following problem description are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
% $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
 $C_i$  & the $i$th cluster in $\mathcal{S}$ \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{theorem}
\label{theorem1}
Robust clustering problem in CRN is NP-hard, when the maximum size of clusters is larger than 3, and $\delta_1=1$ and $\delta_2 = N$.
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.
%This theorem actually indicates that the robust clustering problem in CRN is NP-hard.

We propose a centralized solution which solves an optimization with standard solvers. 
To formulate the optimization, we need to do some preparation beforehand.
All the possible clusters complying with the Description in the system model are found and constitute a set $\mathcal{S}$.
Besides, we assign a weight to each cluster, which correlates with the difference between the cluster size and the desired cluster size. 
In particular, considering $|\mathcal{S}|=M$ and $C_i\in \mathcal{S}$ means the $i$ th cluster in $\mathcal{S}$, and $\delta$ is the desired size, then the weight for each cluster is given as follows,
$$
p_i(C_i) = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$ or $|C_i|=\delta+1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$ or $|C_i|=\delta+2$} \\
\vdots
\end{array} \right.
$$
where $\alpha_1, \alpha_2, \cdots$ are positive values.
In particular, $\alpha$ increases with the divergence between $|C_i|$ and $\delta$ \ie these is $\alpha_2> \alpha_1>0$.

The optimization searches the set $\mathcal{S}$ and decides on certain clusters which constitute the whole CRN without overlapping between any two of them, besides, the sum of CCs of the chosen clusters is maximized.
Then, a central node (or controller) with the knowledge of all CR nodes (also the possible clusters) will solve the problem based on the following formulation.

%Exact cover problem can be solved with Knuth's Algorithm X~\cite{dancingLinks_Knuth} as it finds out all the instances of exact cover, then we can choose the one with the biggest sum of weights. 
%To make the formulation possible, we adopt heuristics in both the objective function and the constraints, besides, the objective is not totally aligned with the object in the Definition~\ref{def_centralized_clustering}.
%The problem by the Definition~\ref{def_centralized_clustering} is formulate

@@

\begin{equation}
\begin{aligned}
     &\min\limits_{w_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (w_i*t_{ij}) \\
     &\text{subject to}   && \Sigma_{i=1}^G x_{ij} = 1, for \forall j=1, \ldots, N \\
   &&& \Sigma_{j=1}^N x_{ij} = |C_i|*w_i, for \forall i=1, \ldots, M \\
   &&& \text{$x_{ij}$ and $w_i$ are binary variables.}\\
   &&& i\in \{1,2, \cdots M\}, \hspace{0.3cm} j\in \{1,2,\cdots N\}
%\notag
\end{aligned}
\label{centralized_opt}
\end{equation}

%The optimization is a binary linear programming problem.
The objective is to maximize the sum of CCs of all the clusters.
%We minimize the opposite of the real objective to make this problem solvable.
$w_i$ and $x_{ij}$ are the two binary variables in this problem.
% where $j\in \{1,2,\cdots,N-1, N\}$ and $i\in \{1,2,\cdots,M-1, M\}$.
$N$ is the total number of CR users in network $\mathcal{N}$, $M = |\mathcal{S}|$.
%$G = |\mathcal{G}|$ and there is $\mathcal{G} \subseteq \mathcal{S}$.
Being either 1 or 0, $w_i$ denotes whether the $i$th cluster $C_i$ in $\mathcal{S}$ is chosen to be in the solution or not.
$x_{ij}$ indicates whether the CR node $j$ resides in the $i$th potential cluster, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$. 
Node index $j$ is identical to the node ID.
%In practice, we can solve the problem with $\mathcal{G} = \mathcal{S}$ when $|\mathcal{S}|$ is not large. 
%
%
%Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{potential} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
%Potential clusters are the clusters which satisfy the conditions introduced in Section~\ref{sec:model}. 
%Note that the potential clusters include the singleton ones.
%By permitting the existence of the singleton clusters, we can ensure that the $\mathcal{S}'$ in Definition~\ref{def_centralized_clustering} is always be feasible.
$t_{ij}$ is a constant which is $p_i(C_i) - \frac{q_{ij}}{|C_i|}$.
$q_{ij}= |K(C_i)|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
$|K(C_i)|$ is the number of CCs of cluster $C_i$, and $|C_i|$ is the size of cluster $C_i$.
%%Based on the knowledge on $\mathcal{S}$, we construct a $M\times N$ matrix $Q_{M\text{x}N}$ which is shown in Figure~\ref{costant_matrix_Q}. 
%%Constant $q_{ij}$ is the element of $Q_{M\text{x}N}$, and its subscripts correspond to the $i$th cluster and CR node $j$ respectively.
%$q_{ij}= |K(C_i)|/|C_i|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
%In other words, each non-zero element $q_{ij}$ means the number of CCs of the cluster $i$ where node $j$ resides.

%\begin{figure}[ht!]
%\centering
%\bordermatrix{~ 		& 1 	& 2 	& 3 	& \cdots & j & \cdots	& N-1 	& N	\cr
%                  1 	& |K(C_1)| 	& |K(C_1)| 	& 0 	& \cdots & \cdots &\cdots	& 0 	& 0	\cr
%                  2 	& |K(C_2)| 	& 0 	& |K(C_2)| 	& \cdots & \cdots & \cdots 	& 0 	& 0	\cr
%				\vdots  	&\vdots & 	 	& 		&  \vdots		& 		& \vdots \cr
%				i 	& 0 	& |K(C_i)| 	& 0 	& \cdots  & \cdots & \cdots 	& |K(C_i)| 	& 0	\cr
%				\vdots  	&\vdots & 	 	& 		&  \vdots & \cdots & \vdots 		& 		& \vdots \cr
%				\vdots 	& 0  	& 0 	& 0 	& \cdots & \cdots & \cdots 	& |K(C_i')| 	& 0	\cr
%				G  	& |K(C_G)| & \cdots	 	& 		&  \vdots	& \cdots & \vdots& 		& \vdots \cr}	
%\caption{An example of Matrix $Q$, its rows correspond to all possible clusters, and columns correspond to the CR nodes in the CRN. }
%\label{costant_matrix_Q}
%\end{figure}

Now we examine the objective function to see whether it in line with the goal to maximize the total number of CCs meanwhile consider the restriction on cluster size.
The objective function can be written as,
\begin{equation}
\begin{aligned}
     &\min\limits_{w_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (-w_i* \frac{q_{ij}}{|C_i|} + w_i*p_i(C_i)) \\
\notag
\end{aligned}
\label{centralized_opt}
\end{equation}
The sum of the first items is the sum of CCs of all the chosen clusters.
The minus sign in front of the first item explains the reason why we minimize instead to maximize the objective function.
As to the second item, when $w_i$ is zero ($C_i$ is chosen), if $|C_i|\neq\delta$, the second component will be positive which contradicts the direction of the optimization.
Thus the second item discourages the appearance the clusters whose sizes are different from $\delta$, especially those whose sizes diverge far from $\delta$.

%$$
%w_i = \left\{ \begin{array}{rl}
%0 &\mbox{if $i$th possible cluster $C_i$ is chosen} \\
%1 &\mbox{if $i$th possible cluster $C_i$ is not chosen} \\
%\end{array} \right.
%$$
%When $w_i$ is zero, it means $C_i$ is chosen by the optimization algorithm, in this case, the second item in the objective function becomes $p$, which is defined as follows,

%Because of $w_i$, any cluster which appears in the clustering solution ($w=0$) will compromise the first component.
%Only when that cluster's size is exactly the preferred size $\delta$, the punishment is zero.
%In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
%Then it is easy to notice that when the optimization algorithm adopts singleton clusters, the first xx.
%With such design, the solution is encouraged not to generate singleton clusters.
%doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
%Choice of $\alpha_i$ affects the resultant clusters.
